Theories of deliberated choice

### 2023-11-10 simple.tex
Cleaning fals

That the deliberated choice of any individual $i$ is well-defined may be doubtful, as it involves counterfactuals.
To see this, let $\hist = (a_1, a_2, a_e)$ designate the argumentative path where the arguments are shown in the order they are described in \cref{ex:tutor}, and let $\hist' = (a_e, a_2, a_1)$ designate the reverse order.
The deliberated choice concept is defined so as to not depend on the order of the arguments shown to $i$: for having $\prop \in P_i$, it must be that directing to $i$ the path $\hist$ or the path $\hist'$ both lead $i$ to choose $\prop$.
However, once we have shown $\hist$ to $i$, we cannot show to her $\hist'$ (unless assuming $i$ forgets arguments after a while and waiting for sufficiently long, which would create practical and theoretical difficulties).

This difficulty is resolved using falsifiable general theories, applicable to $I$ and not solely to a given $i$, so that $\hist$ can be tested on $i$ and $\hist'$ on some $i' ≠ i$. Indeed, we need a way to empirically test claims such as the one concluding \cref{ex:tutor}, according to which $a_e$ is a decisive argument. This is achieved in the next section thanks to theories of deliberated choice.
% Such theories contain an observable claim and a conclusion claim, so as to satisfy two properties. First, such a theory must be falsifiable in principle, meaning that it must be possible, using our observational basis (propositions of the sort $\alpha \ileadsto \phi$), to invalidate the observable claim of the theory. As a counter-example, using \cref{ex:tutor}, a theory that claims that $\forall i \in I: \alpha_{123} \ileadsto \prop_1 \lor \alpha_{321} \ileadsto \prop_1$ is not falsifiable in principle, as, whatever the real $\ileadsto$ relation, with a given $i$, one can never show that $\alpha_{123} \nileadsto \prop_1 \land \alpha_{123} \nileadsto \prop_1$. Second, such a theory must be so that if its observable claim holds, then some given alternative is in the deliberated judgment of the individuals in $I$.

\subsection{Form of theories}
Given an exposure protocol $(I, \allargs, {\allleadsto})$, recall that ${\allleadsto} \in \allleadstos$.

Given a pair $(I, \allargs)$, a claim $C$ is a set of possibilities for $\allleadsto$, thus, $C \subseteq \allleadstos$.
Given a formula $\psi$, let the corresponding claim be $C_\psi = \set{{\allleadstog} \in \allleadstos \suchthat {\allleadstog} \text{ satisfies } \psi}$.
A claim is trivial iff it is the trivial claim, defined as $C = \allleadstos$.
A claim is finite iff its cardinal is finite (meaning that it is compatible with a finite number of possibilities).
Let $C^\subseteq = \set{c \subseteq (I × \allhists) × P \suchthat \exists {\allleadstog} \in C \suchthat c \subseteq {\allleadstog}}$ denote the set of non necessarily complete possibilities compatible with $C$ (thus, each element $c \in C^\subseteq$ is right-unique but non necessarily total, meaning that it does not necessarily cover $(I × \allhists)$).
Note that $C^\subseteq$ is downwards closed, meaning that $\forall c \in C^\subseteq: c' \subseteq c \implies c' \in C^\subseteq$.

Given a pair $(I, \allargs)$, an observation $θ \in \Theta$ is a right unique finite subset of $(I × \allhists) × P$ such that $\forall i \in I, (i, \hist, p), (i, \histp, q) \in θ: \card{\hist} ≤ \card{\hist'} \implies \hist \subseteq \hist'$.
The set $\Theta$ contains the a priori possible observations, that is, those that can be considered possible when nothing is known about $\allleadsto$.
Let $S_θ = \set{{\allleadstog} \in \allleadstos \suchthat θ \subseteq {\allleadstog}}$ denote the possibilities for $\allleadsto$ compatible with the observations.
We say that an observation $\theta$ is compatible with a claim $C$ iff $S_\theta \cap C ≠ \emptyset$, equivalently, iff $\theta \in C^\subseteq$.

@@

\begin{remark}
  Any relation $\allleadstog$ is equivalent to $\cup \set{\restr{\allleadstog}{O} \suchthat O \in \cup_{J \subseteq_\finite I} (\allhists)^J}$.
\end{remark}

\begin{definition}
  We say that $C$ is \emph{syntactically falsifiable}, or just falsifiable, iff
  %$\forall {\allleadstog} \in \allleadstos: ¬C({\allleadstog}) \implies \exists θ \subseteq {\allleadstog} \suchthat θ \implies ¬C$, meaning that
  $\forall {\allleadstog} \notin C: \exists θ \subseteq {\allleadstog} \suchthat S_θ \cap C = \emptyset$.
\end{definition}
Equivalently, iff $\forall {\allleadstog} \notin C: \exists θ \subseteq {\allleadstog} \suchthat \theta \notin C^\subseteq$.
Equivalently, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \forall \theta \in \Theta, \theta \subseteq {\allleadstog}: \theta \in C^\subseteq$.

\begin{example}
  \label{ex:bigor}
  $[\forall i: (\ar) \ileadstog \phi] \lor [\forall i: (\ar, \ar) \ileadstog \phi] \lor …$ is not falsifiable.
  Here, $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$.

  Given any $θ$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in θ} \card{\hist} + 1$ if $\exists p \in P \suchthat (i, \hist, p) \in θ$, and $l = 1$ otherwise. Pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in S_θ \cap C ≠ \emptyset$.
\end{example}
\begin{example}
  \label{ex:disj}
  Given $\ar \neq \ar' \in \allargs$,
  $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
  Consider any ${\allleadstog} \notin C$, thus satisfying, for some $i$, $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
  Consider any $θ \subseteq {\allleadstog}$ and let us prove that $S_θ \cap C ≠ \emptyset$.
  By definition of a general observation, $θ$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $θ$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in S_θ \cap C ≠ \emptyset$.
\end{example}
\begin{example}
  \label{ex:partlyU}
  Given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
  $\hist \ileadstog[j] \phi \land \forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
  Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
  \Cref{ex:disj} concludes.
\end{example}
\begin{example}
  \label{ex:allParticular}
  Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
  Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $θ = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
  \label{ex:existsParticular}
  Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
  Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $θ \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $S_θ \cap C ≠ \emptyset$.
\end{example}

@@

\begin{theorem}
  \label{thm:caractFals}
  A claim $C$ is falsifiable iff $\forall {\allleadstog} \in \allleadstos$:
  \[\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq \implies {\allleadstog} \in C.\]
\end{theorem}
\begin{proof}
  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: S_θ \cap C \neq \emptyset$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: θ \in C^\subseteq$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.

@@

We thus reach a procedure to determine whether a claim is falsifiable.
Given $C$, compute $C^\subseteq$. Find the set $R$ of relations $\allleadstog$ satisfying $\forall O: \restr{\allleadstog}{O} \in C^\subseteq$. The claim $C$ is falsifiable iff $R \subseteq C$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq} \subseteq C$.
\end{corollary}
A relation $\allleadstog$ is observably in $C$ iff $\forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq$, equivalently, iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)$.
All relations in $C$ are observably in $C$. The question of falsifiability asks whether the converse holds.
In other words, $C$ is falsifiable iff it includes the relations observably in $C$.
Define that a relation ${\allleadstog} \in {\allleadstos}$ is \emph{obs-compatible} with a set $S$ iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S$.
Then, the relations observably in $C$ are exactly those obs-compatible with $C^\subseteq \cap \Theta$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)} = C$.
\end{corollary}
\begin{remark}
  The question of falsifiability is whether the set of relations $\allleadstog$ “compatible” with $C^\subseteq \cap \Theta$ equals $C$.
  Compatibility needs to consider the observables, not a subset of it.
  Define $\Theta_1 \subset \Theta$ as the observations that observe only one pair $(i, \hist)$.
  Define $R_1 = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta_1: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)}$.
  Note that $R \subseteq R_1$.
  We know that $C$ is falsifiable iff $R \subseteq C$. Thus, $C$ is not falsifiable implies $C \subsetneq R \subseteq R_1$ thus $C \subsetneq R_1$. But we do not have that $C \subsetneq R_1$ implies $C$ not falsifiable: could have $C \subsetneq R_1$ but still $R = C$.
\end{remark}
\begin{example}
  Considering again \cref{ex:bigor}, by \cref{th:veryunfals}, we see that $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$ is not falsifiable.

@@

  Pick some $o_1 \in R$ as in the hypothesis.
  Consider any $o \in O_1 \cup \set{o_2}$.
  As $\set{o_1, o} \in \allpriobss$ and ${\allleadstog}$ being obs-compatible, we have
  $\restr{\allleadstog}{\set{o_1, o}} \in C^\wsub$.
  As also $o_1, o \in R$, ${\allleadstog}(o) = f(o)$.
  In particular, ${\allleadstog}(o_2) = f(o_2)$.
  Iterate with $o_2$ and $o \in O_2$, and so on.
  We obtain that $\forall o \in R: {\allleadstog}(o) = f(o)$, therefore, ${\allleadstog} \in C$.
\end{proof}
\begin{example}
  Pick $i \in I$ and
  define $o_1 = ((a), i)$,
  $o_2 = (\emptyset, i)$ and
  $o_3 = ((b), i)$.
  Define $R = \set{o_1, o_2, o_3}$.
  Define $f_1, f_2, f_3 \in P^R$ as follows.
  $f_1(o_1) = \phi$,
  $f_1(o_2) = ¬\phi$,
  $f_1(o_3) = \phi$.
  $f_2(o_1) = \phi$,
  $f_2(o_2) = \phi$,
  $f_2(o_3) = ¬\phi$.
  $f_3(o_1) = ¬\phi$,
  $f_3(o_2) = \phi$,
  $f_3(o_3) = \phi$.
  Define $C = \set{f \cup g \suchthat f \in \set{f_1, f_2, f_3} \land g \in P^{(I × \allhists) \setminus R}}$.

@@

@@

Given an exposure protocol $\allleadsto$, we say that a claim $C$ \emph{holds} iff $C \ni {\allleadsto}$.
We say that $C$ is \emph{empirically falsifiable} iff $\exists θ \subseteq {\allleadsto} \suchthat S_θ \cap C = \emptyset$.

The following theorem verifies that when $C$ holds, it is not empirically falsifiable (though it may be syntactically falsifiable).
\begin{theorem}
  \label{th:ex-post}
  $\forall C \ni {\allleadsto}: \nexists θ \subseteq {\allleadsto} \suchthat S_θ \cap C = \emptyset$.
\end{theorem}
\begin{proof}
  We have, $\forall C \suchthat C \ni {\allleadsto}, \forall θ \subseteq {\allleadsto}$: $θ \subseteq {\allleadsto} \in C$, equivalently, ${\allleadsto} \in S_θ \cap C$, thus $S_θ \cap C ≠ \emptyset$.
\end{proof}
Therefore, synctactic falsifiability guarantees that a claim holds if, and only if, it is not empirically falsifiable.
\begin{theorem}
  \label{th:holdsiff}
  For all syntactically falsifiable $C$: ${\allleadsto} \in C$ iff $\nexists θ \subseteq {\allleadsto} \suchthat S_θ \cap C = \emptyset$.
\end{theorem}
\begin{proof}
  Consider any syntactically falsifiable $C$.
  By definition of falsifiability, $\forall {\allleadstog}: \nexists θ \subseteq {\allleadstog} \suchthat S_θ \cap C = \emptyset \implies {\allleadstog} \in C$.
  In particular,
  $\nexists θ \subseteq {\allleadsto} \suchthat S_θ \cap C = \emptyset \implies {\allleadsto} \in C$.
  The reverse implication is proven by \cref{th:ex-post}.
\end{proof}

>>>>

Now that deliberated choices are defined, we need to explore what one can learn about deliberated choices using our empirically available data.
This requires to define claims about deliberated choices and to explore how to validate claims.

\section{Claims about deliberated choice}
Given an exposure protocol $(I, \allargs, {\allleadsto})$, consider a claim about $\allleadsto$, such as “every individuals in $I$ choose deliberately $\phi$”, or, in \cref{ex:tutor}, “$a_e$ is a decisive argument for $i$ in favor of $\phi_2$”.
Validating claims about deliberated choice is no trivial matter, for two reasons.

First,
such claims must entail an infinity of constraints about $\allleadsto$:
in order to deduce that $p \in P_i$ for some $p \in P, i \in I$, the claim must constrain $\ileadsto$ on every possible paths $\alpha \in \allhists$.
It follows that a claim cannot be validated empirically exhaustively by observing the value of $\allleadsto$ on every points that it constraints.

Second,
our observations are limited in important ways: once we have shown some arguments to a given individual $i$, we cannot undo the effects of this action on $i$ and show to her a different argument as if starting from scratch (unless assuming $i$ forgets arguments after a while and waiting for sufficiently long, which would create practical and theoretical difficulties).
In other words, once we have shown a path $\hist$ to $i$, we can only continue the protocol with $i$ by showing paths $\histp$ that are suffixes of $\hist$.

To illustrate this point with \cref{ex:tutor}, let $\hist = (a_1, a_2, a_e)$ designate the argumentative path where the arguments are shown in the order they are described in \cref{ex:tutor}, and let $\hist' = (a_e, a_2, a_1)$ designate the reverse order.
The deliberated choice concept is defined so as to not depend on the order of the arguments shown to $i$: for having $\prop \in P_i$, it must be that directing to $i$ the path $\hist$ or the path $\hist'$ both lead $i$ to choose $\prop$.
But it is impossible to observe both the value of ${\allleadsto}(i, \hist)$ and the value of ${\allleadsto}(i, \hist')$.

The first difficulty is familiar as it occurs in any endeavor to validate general claims in science: claims that cover a large (or infinite) spectrum cannot be validated by observing the system exhaustively. One must rather develop a falsification protocol, that is, a way to invalidate the claim, and then observe that the claim is not invalidated. This is the approach taken in the next section.

This approach also permits to tackle the second difficulty, by letting us study theoretically the needs and limits of falsification given our observable basis.
% Such theories contain an observable claim and a conclusion claim, so as to satisfy two properties. First, such a theory must be falsifiable in principle, meaning that it must be possible, using our observational basis (propositions of the sort $\alpha \ileadsto \phi$), to invalidate the observable claim of the theory. As a counter-example, using \cref{ex:tutor}, a theory that claims that $\forall i \in I: \alpha_{123} \ileadsto \prop_1 \lor \alpha_{321} \ileadsto \prop_1$ is not falsifiable in principle, as, whatever the real $\ileadsto$ relation, with a given $i$, one can never show that $\alpha_{123} \nileadsto \prop_1 \land \alpha_{123} \nileadsto \prop_1$. Second, such a theory must be so that if its observable claim holds, then some given alternative is in the deliberated judgment of the individuals in $I$.

\subsection{Claims and observable basis}
% Such a deduction should use, as much as possible, the observational basis, that is, the propositions of the form $\alpha \ileadsto p$ that can be obtained experimentally.
Note first that given a relation ${\allleadstog} \in \allleadstos$, $\powerset{\allleadstog}$ denotes the set of relations over $(I × \allhists) × P$ that refine $\allleadstog$ (thus, each element $f \in \powerset{\allleadstog}$ is right-unique but non necessarily total, meaning that it does not necessarily cover $(I × \allhists)$).

Given a pair $(I, \allargs)$, a claim $C$ is a set of possibilities for $\allleadsto$, thus, $C \subseteq \allleadstos$.
Given a formula $\psi$, let the corresponding claim be $C_\psi = \set{{\allleadstog} \in \allleadstos \suchthat {\allleadstog} \text{ satisfies } \psi}$.
A claim is trivial iff it is the trivial claim, defined as $C = \allleadstos$.
A claim is finite iff its cardinal is finite (meaning that it is compatible with a finite number of possibilities).
Let $C^\wsub = \cup_{{\allleadstog} \in C} \powerset{\allleadstog} = \set{c \subseteq (I × \allhists) × P \suchthat \exists {\allleadstog} \in C \suchthat c \in \powerset{\allleadstog}}$ denote the set of non necessarily complete possibilities compatible with $C$ (thus, each element $c \in C^\subseteq$ is right-unique but non necessarily total, meaning that it does not necessarily cover $(I × \allhists)$).
Note that $C^\subseteq$ is downwards closed, meaning that $\forall c \in C^\subseteq: c' \subseteq c \implies c' \in C^\subseteq$.

Given a pair $(I, \allargs)$, an observation $θ \in \Theta$ is a right unique finite subset of $(I × \allhists) × P$ such that $\forall i \in I, (i, \hist, p), (i, \histp, q) \in θ: \card{\hist} ≤ \card{\hist'} \implies \hist \subseteq \hist'$.
The set $\Theta$ contains the a priori possible observations, that is, those that can be considered possible when nothing is known about $\allleadsto$.
% Let $S_θ = \set{{\allleadstog} \in \allleadstos \suchthat θ \subseteq {\allleadstog}}$ denote the possibilities for $\allleadsto$ compatible with the observations.
We say that an observation $\theta$ is compatible with a claim $C$ iff
% $S_\theta \cap C ≠ \emptyset$, equivalently, iff 
$\theta \in C^\subseteq$, equivalently, iff
$\theta \in \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
Thus, given ${\allleadstog} \in \allleadstos$, the set $\Theta \cap \powerset{\allleadstog}$ is the set of observations compatible with ${\allleadstog}$.

@@

\subsection{Rejection and falsification}
Given an exposure protocol $(I, \allargs, {\allleadsto})$, and a claim about ${\allleadsto}$, we develop here a procedure to deduce that the claim is correct.

Given an exposure protocol $\allleadsto$, we say that a claim $C$ \emph{holds} iff $C \ni {\allleadsto}$.

At our disposal is the procedure of empirical rejection, that consists in directly invalidating a claim by observing something that contradicts it.
\begin{definition}
  Given an exposure protocol $\allleadsto$, we say that $C$ is \emph{rejectable} iff
  % $\exists θ \in \Theta \cap \powerset{\allleadsto} \suchthat θ \notin C^\wsub$.
  $\Theta \cap \powerset{\allleadsto} \nsubseteq C^\wsub$.
\end{definition}

The following theorem verifies that when $C$ holds, it is not rejectable.
\begin{theorem}
  \label{th:ex-post}
  $\forall C \ni {\allleadsto}:
    % \nexists θ \in \Theta \suchthat \theta \subseteq {\allleadsto} \land θ \notin C^\wsub$.
  \Theta \cap \powerset{\allleadsto} \subseteq C^\wsub$.
\end{theorem}
\begin{proof}
  This holds by definition of $C^\wsub = \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
  % Consider any claim $C$ with ${\allleadsto} \in C$.
  % Consider any $\theta \in \Theta$ with $\theta \wsub {\allleadsto}$.
  % We have $θ \subseteq {\allleadsto} \in C$, thus $θ \in C^\wsub$.
\end{proof}

The converse, however, does not hold.
\begin{example}[Disjunct with incompatible paths, not rejectable]
  \label{ex:disjR}
  Given $\ar \neq \ar' \in \allargs$,
  the claim $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not rejectable.
  In fact, 
  % $\nexists \theta \in \Theta \suchthat \theta \notin C^\wsub$.
  $\Theta \wsub C^\wsub$.
  % Consider any ${\allleadstog} \notin C$, thus satisfying, for some $i$, $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
  Consider any $θ \in \Theta$ and let us prove that $\theta \in \cup_{{\allleadstog} \in C} \powerset{\allleadstog}$.
  By definition of $\Theta$, $\theta^{-1}(P)$ may not contain both $((\ar), i)$ and $((\ar'), i)$, thus letting $t$ denote one of these pairs not in $\theta^{-1}(P)$ and defining $r = (t, \phi)$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have $\theta \in \powerset{\allleadstog'}$ and ${\allleadstog'} \in C$.
\end{example}
\Cref{ex:disjR} shows a claim that is not rejectable, irrespective of the exposure protocol.
In particular, the example applies when $C$ does not hold (that is, when ${\allleadsto} \notin C$). We conclude that a claim may fail to hold but still be not rejectable.
It follows that non-rejection cannot be a sufficient condition for validating a claim.

\begin{remark}
  \Cref{ex:threeFs} provides an example where, similarly, the claim may fail to hold but still be not rejectable; but where rejection depends on ${\allleadsto}$.
\end{remark}

\Cref{def:fals} fills the gap to obtain that $C$ holds iff it is not rejectable.
The property demands that whatever the attitude of individuals, if the attitude is not according to the claim, then it is possible to observe that it is not.
\begin{definition}
  \label{def:fals}
  A claim $C$ is \emph{falsifiable} iff
  %$\forall {\allleadstog} \in \allleadstos: ¬C({\allleadstog}) \implies \exists θ \subseteq {\allleadstog} \suchthat θ \implies ¬C$, meaning that
  $\forall {\allleadstog} \notin C: \exists θ \in \Theta \cap \powerset{\allleadstog} \suchthat \theta \notin C^\wsub$.
\end{definition}
% A claim $C$ is \emph{falsifiable} iff $\forall {\allleadstog} \notin C: \exists θ \subseteq {\allleadstog} \suchthat \theta \notin C^\subseteq$,
% equivalently iff $\exists {\allleadstog} \notin C \suchthat \forall \theta \in \Theta, \theta \subseteq {\allleadstog}: \theta \in C^\subseteq$.

It is important to note that falsifiability does not depend on $\allleadsto$: whether falsifiability holds can be determined analytically, meaning, by considering only the content of the claim, without considering the behavior of individuals.

\Cref{th:holdsiff} shows that falsifiability indeed fills that gap:
falsifiability guarantees that a claim holds if, and only if, it is not rejectable.
\begin{theorem}
  \label{th:holdsiff}
  For any falsifiable claim $C$: $C \ni {\allleadsto}$ iff $\Theta \cap \powerset{\allleadsto} \wsub C^\wsub$.
\end{theorem}
\begin{proof}
  Consider any falsifiable $C$.
  Using the contrapositive of the definition of falsifiability, $\forall {\allleadstog} \in \allleadstos: \nexists θ \in \Theta \cap \powerset{\allleadstog} \suchthat \theta \notin C^\wsub \implies {\allleadstog} \in C$.
  In particular,
  $\nexists θ \in \Theta \cap \powerset{\allleadsto} \suchthat \theta \notin C^\wsub \implies {\allleadsto} \in C$.
  The reverse implication is proven by \cref{th:ex-post}.
\end{proof}

\begin{example}[Big disjunct, not falsifiable]
  \label{ex:bigor}
  $[\forall i: (\ar) \ileadstog \phi] \lor [\forall i: (\ar, \ar) \ileadstog \phi] \lor …$ is not falsifiable.
  Here, $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$.

  Given any $θ$, define $l = \max_{\hist \suchthat \exists p \in P \suchthat (i, \hist, p) \in θ} \card{\hist} + 1$ if $\exists p \in P \suchthat (i, \hist, p) \in θ$, and $l = 1$ otherwise. Pick any $\allleadstog$ such that $θ \subset {\allleadstog} \land (\ar)^l \ileadstog \phi$. We have ${\allleadstog} \in \theta \in C^\wsub$.
\end{example}
\begin{example}[Disjunct with incompatible paths, not falsifiable]
  \label{ex:disj}
  Given $\ar \neq \ar' \in \allargs$, the claim appearing in \cref{ex:disjR}, namely,
  $\forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$, is not falsifiable.
  Consider any ${\allleadstog} \notin C$, thus satisfying, for some $i$, $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$.
  Consider any $θ \subseteq {\allleadstog}$ and let us prove that $\theta \in C^\wsub$.
  By definition of a general observation, $θ$ may not contain both $((\ar), i, \phi)$ and $((\ar'), i, \phi)$, thus letting $t$ denote one of these triples not in $θ$, pick any ${\allleadstog'}$ such that $θ \subset {\allleadstog'} \land t \in {\allleadstog'}$. We have ${\allleadstog'} \in \theta \in C^\wsub$.
\end{example}
\begin{example}[Partly falsifiable is still unfalsifiable]
  \label{ex:partlyU}
  Given $j \in I$, $\hist \in \allhists$ and $\ar \neq \ar' \in \allargs$,
  $\hist \ileadstog[j] \phi \land \forall i: [(\ar) \ileadstog \phi \lor (\ar') \ileadstog \phi]$ is not falsifiable.
  Consider any $\allleadstog$ satisfying $(\ar) \nileadstog \phi \land (\ar') \nileadstog \phi$, thus with ${\allleadstog} \notin C$.
  \Cref{ex:disj} concludes.
\end{example}
\begin{example}
  \label{ex:allParticular}
  Consider any $\ar$ and $\phi$. The claim $\forall i: \ar \ileadstostg \phi$ is falsifiable.
  Given any ${\allleadstog} \notin C$, consider any $i \suchthat \ar \nileadstostg \phi$ and any $\hist \suchthat \hist \nileadstog \phi \land \ar \in \histend$, and define $θ = \set{(i, \hist, {\allleadstog}(i, \hist) ≠ \phi)}$.
\end{example}
\begin{example}
  \label{ex:existsParticular}
  Given an infinite $I$, the claim $\exists i \suchthat \ar \ileadstostg \phi$ is not falsifiable.
  Given any ${\allleadstog} \notin C$, thus such that $\forall i: \ar \nileadstostg \phi$, any $θ \subseteq {\allleadstog}$ only has finitely many $(i, \hist, p ≠ \phi)$ thus $\theta \in C^\wsub$.
\end{example}

@@

\begin{theorem}
  \label{thm:caractFals}
  A claim $C$ is falsifiable iff $\forall {\allleadstog} \in \allleadstos$:
  \[\set{\restr{\allleadstog}{O} \suchthat O \in \allpriobss} \subseteq C^\subseteq \implies {\allleadstog} \in C.\]
\end{theorem}
\begin{proof}
  By definition, $C$ is not falsifiable iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: \theta \in C^\wsub$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \forall θ \subseteq {\allleadstog}: θ \in C^\subseteq$;
  equivalently, iff $\exists {\allleadstog} \notin C \suchthat \set{θ \subseteq {\allleadstog}} \subseteq C^\subseteq$.

@@

We thus reach a procedure to determine whether a claim is falsifiable.
Given $C$, compute $C^\subseteq$. Find the set $R$ of relations $\allleadstog$ satisfying $\forall O: \restr{\allleadstog}{O} \in C^\subseteq$. The claim $C$ is falsifiable iff $R \subseteq C$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq} \subseteq C$.
\end{corollary}
A relation $\allleadstog$ is observably in $C$ iff $\forall O \in \allpriobss: \restr{\allleadstog}{O} \in C^\subseteq$, equivalently, iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)$.
All relations in $C$ are observably in $C$. The question of falsifiability asks whether the converse holds.
In other words, $C$ is falsifiable iff it includes the relations observably in $C$.
Define that a relation ${\allleadstog} \in {\allleadstos}$ is \emph{obs-compatible} with a set $S$ iff $\forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in S$.
Then, the relations observably in $C$ are exactly those obs-compatible with $C^\subseteq \cap \Theta$.
\begin{corollary}
  A claim $C$ is falsifiable iff $\set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)} = C$.
\end{corollary}
\begin{remark}
  The question of falsifiability is whether the set of relations $\allleadstog$ “compatible” with $C^\subseteq \cap \Theta$ equals $C$.
  Compatibility needs to consider the observables, not a subset of it.
  Define $\Theta_1 \subset \Theta$ as the observations that observe only one pair $(i, \hist)$.
  Define $R_1 = \set{{\allleadstog} \in \allleadstos \suchthat \forall \theta \in \Theta_1: ({\allleadstog} \cap \theta) \in (C^\subseteq \cap \Theta)}$.
  Note that $R \subseteq R_1$.
  We know that $C$ is falsifiable iff $R \subseteq C$. Thus, $C$ is not falsifiable implies $C \subsetneq R \subseteq R_1$ thus $C \subsetneq R_1$. But we do not have that $C \subsetneq R_1$ implies $C$ not falsifiable: could have $C \subsetneq R_1$ but still $R = C$.
\end{remark}
\begin{example}[Come back]
  Considering again \cref{ex:bigor}, by \cref{th:veryunfals}, we see that $C = \set{{\allleadstog} \suchthat \exists n \in \N^* \suchthat \forall i: (a)^n \ileadstog \phi}$ is not falsifiable.

@@

  Pick some $o_1 \in R$ as in the hypothesis.
  Consider any $o \in O_1 \cup \set{o_2}$.
  As $\set{o_1, o} \in \allpriobss$ and ${\allleadstog}$ being obs-compatible, we have
  $\restr{\allleadstog}{\set{o_1, o}} \in C^\wsub$.
  As also $o_1, o \in R$, ${\allleadstog}(o) = f(o)$.
  In particular, ${\allleadstog}(o_2) = f(o_2)$.
  Iterate with $o_2$ and $o \in O_2$, and so on.
  We obtain that $\forall o \in R: {\allleadstog}(o) = f(o)$, therefore, ${\allleadstog} \in C$.
\end{proof}
\begin{example}
  Pick $i \in I$ and
  define $o_1 = ((a), i)$,
  $o_2 = (\emptyset, i)$ and
  $o_3 = ((b), i)$.
  Define $R = \set{o_1, o_2, o_3}$.
  Define $f_1, f_2, f_3 \in P^R$ as follows.
  $f_1(o_1) = \phi$,
  $f_1(o_2) = ¬\phi$,
  $f_1(o_3) = \phi$.
  $f_2(o_1) = \phi$,
  $f_2(o_2) = \phi$,
  $f_2(o_3) = ¬\phi$.
  $f_3(o_1) = ¬\phi$,
  $f_3(o_2) = \phi$,
  $f_3(o_3) = \phi$.
  Define $C = \set{f \cup g \suchthat f \in \set{f_1, f_2, f_3} \land g \in P^{(I × \allhists) \setminus R}}$.
  This is the claim of \cref{ex:threeFs}.

@@

  Note that the complement is as follows.
  \begin{equation}
    \overline{C^\wsub \cap \Theta} = \left\{\theta \in \Theta \suchthat
    \begin{aligned}
       & \theta(o_2) = 0 \lor \mbox{}                                                                                \\
       & \theta(o_2) = ¬\phi \land (\theta(o_1) \in \set{¬\phi, 0} \lor \theta(o_3) \in \set{¬\phi, 0}) \lor \mbox{} \\
       & \theta(o_1) \in \set{¬\phi, 0} \land \theta(o_3) \neq ¬\phi \lor \mbox{}                                    \\
       & \theta(o_1) = \emptyset \lor \mbox{}                                                                        \\
       & \theta(o_3) = \emptyset \lor \mbox{}                                                                        \\
       & \theta(o_3) \in \set{¬\phi, 0} \land \theta(o_1) \neq ¬\phi
    \end{aligned}\right\}.
  \end{equation}
  This seems incorrect: if $\theta = \emptyset$, then it belongs, even though it satisfies $\theta(o_1) = \emptyset$.

  The relations that are obs-compatible with $C^\wsub \cap \Theta$ are the ones that, for $(o_1, o_2, o_3)$, include two $\phi$ and one $¬\phi$, the one that equals all $\phi$, and the one that assigns $o_1$ and $o_3$ to $¬\phi$ and $o_2$ to $\phi$.

  It follows that the relations that are obs-compatible with $C^\wsub \cap \Theta$ and that are in $C^\wsub \cap \Theta$ are the ones that, for $(o_1, o_2, o_3)$, include two $\phi$ and one $¬\phi$; and the relations that are obs-compatible with $C^\wsub \cap \Theta$ and that are not in $C^\wsub \cap \Theta$ are the one that equals all $\phi$ and the one that assign $o_1$ and $o_3$ to $\emptyset$ and $o_2$ to $\phi$.

@@
